---
title: "p8105_hw2_yc4587"
author:"Yifei Chen"
date:"2024-10-2"
output: github_document
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The cleaned dataset now includes the following variables:

line: The subway line (e.g., "1", "2", "A")
station_name: The name of the station
station_latitude and station_longitude: Geographical coordinates of the station
routes: Columns representing different subway routes served by the station
entry: A logical variable (TRUE if the entrance allows entry, FALSE otherwise)
vending: Whether the station has vending machines (e.g., "YES" or "NO")
entrance_type: The type of entrance (e.g., "Stair", "Elevator")
ada: Whether the station is ADA compliant (e.g., "YES" or "NO")
 

```{r}
trans_ent |> 
  select(station_name, line) |> 
  distinct()
```

We filter the data where ada is "true" and count the number of rows.

```{r}
trans_ent |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

We filter the data where vending == "NO" and calculate the proportion of entrances where entry == TRUE.

```{r}
trans_ent |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant.

```{r}
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()

trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
The dataset provides details about subway entrances and exits across NYC, including routes, ADA compliance, and entry types.
After cleaning the data, there are 468 distinct stations, 107 stations are ADA compliant. About 37.6% of stations without vending machines allow entry. The A train serves 60 distinct stations, and 17 of those stations are ADA compliant.

### Problem 2 

```{r}
library(readxl)
library(tidyverse)

# Read and clean Mr. Trash Wheel data
mr_trash_wheel <- read_excel("data/202409 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", skip = 1) %>%
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>%  # Keep rows with dumpster-specific data
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    year = as.integer(year)
  )

```

We skip non-data rows: The skip = 1 argument skips the first row, which contains non-data entries. Omit rows without dumpster-specific data: The filter(!is.na(dumpster)) removes rows where dumpster is NA (i.e., rows that don’t contain data for a specific dumpster). The janitor::clean_names() function is used to clean and standardize column names, removing spaces and converting them to snake_case. We round and convert sports balls that the mutate() function rounds the number of sports balls to the nearest integer and converts the result to an integer using as.integer().

```{r}
# Read and clean Professor Trash Wheel data
prof_trash_wheel <- read_excel("data/202409 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", skip = 1) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(person = "prof_trash_wheel", 
         year = as.integer(year))

# Read and clean Gwynnda Trash Wheel data
gwynnda_trash_wheel <- read_excel("data/202409 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", skip = 1) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(person = "gwynnda_trash_wheel", 
         year = as.integer(year))
```

```{r combine dataset}
# Combine the three datasets

combined_trash_wheel_data <- bind_rows(
  mr_trash_wheel,
  prof_trash_wheel,
  gwynnda_trash_wheel
)
```

a) What is the total weight of trash collected by Professor Trash Wheel?
We can filter the combined dataset for "Professor Trash Wheel" and sum the total weight of trash.
```{r}
# Total weight of trash collected by Professor Trash Wheel
total_weight_professor_trash_wheel <- combined_trash_wheel_data %>%
  filter(person == "prof_trash_wheel") %>%
  summarise(total_weight = sum(weight_tons, na.rm = TRUE))

print(total_weight_professor_trash_wheel)
```

b) What was the total number of cigarette butts collected by Gwynnda in June 2022?
For this, filter the dataset for Gwynnda, select rows corresponding to June 2022, and sum the total number of cigarette butts.
```{r}
# Total cigarette butts collected by Gwynnda in June 2022
total_cigarette_gwynnda_trash_wheel <- combined_trash_wheel_data %>%
  filter(person == "gwynnda_trash_wheel", month == "June", year == "2022") %>%
  summarise(total_number_cigarette_butts = sum(cigarette_butts, na.rm = TRUE))

print(total_cigarette_gwynnda_trash_wheel)
```
Professor Trash Wheel collected a total of 246.74 tons of trash, while Gwynnda collected 18120 cigarette butts in June 2022.

### Problem 3 

library(dplyr)
library(readr)

setwd("/Users/sofiachen/Desktop/p8105_hw2_yc4587/data")

bakers <- read_csv("bakers.csv")
bakes <- read_csv("bakes.csv")
results <- read_csv("results.csv", skip = 2)
viewers <- read_csv("viewers.csv")

bakes <- bakes %>% rename("series" = "Series", "episode" = "Episode", "baker" = "Baker")
bakers <- bakers %>% rename("series" = "Series")
bakers <- bakers %>% mutate(baker = sub(" .*", "", `Baker Name`))

result_bake <- merge(results, bakes, by = c("series", "episode", "baker"))
final_result <- merge(result_bake, bakers, by = c("series", "baker"))

final_result <- final_result %>% mutate(rank = factor(result, levels = c("WINNER", "Runner-up", "IN", "OUT")))

final_result <- final_result %>% arrange(series, episode, rank)

final_result <- final_result %>% select(-baker, -rank)

write.csv(final_result, "final_result.csv", row.names = FALSE)

# We deal with missing data by removing rows with missing values and ensuring consistency of data types. We merge the roaster, bake-off and results data sets into one data set that includes roasters, bake-offs and their performance. The list of star roasters or winners from Seasons 5 through 10 will show predictable patterns or surprises in the competition. In addition, we calculated average ratings for Season 1 and Season 5 to track the popularity of the show.







